Probability

### 2016-04-19 prob.tex
From: Tai, Yu-Tsung Sent: Monday, April 18, 2016 23:55 To: Sabry, Amr A. <sabry@indiana.edu>; Hanson, Andrew J. <hansona@indiana.edu>; Ortiz, Gerardo <ortizg@indiana.edu> Subject: RE: probability in the quantum case
Dear Prof. Sabry, 

I have updated the plan section, and I updated my comment so that it may
be clearer than previously... We may meet and discuss if it is still not
clear enough...

About Prof. Ellerman’s paper, it seems interested, although his bra and
the probability he defined is basis dependent… I would be really
interested if he could build a probability model on a basis inde

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbm}
\usepackage{bbold}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{array}
\usepackage{mathrsfs}

@@

We now formalize a \emph{quantum probability space} as
follows~\cite{Maassen2010,Swart2013}. We first assume an ambient
Hilbert space~$\Hilb$ and define the set of events $\qevents$ as all
linear combinations of \emph{projections} on~$\Hilb$. Specifically,
let $P_1, P_2, \ldots, P_k$ be mutually orthogonal projections on
$\Hilb$ with sum $\mathbb{1}$ and define the event space $\qevents$ to
be the linear span of these operators:
\[
\qevents=\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} .
\]
Each quantum state $\ket{\psi}$ induces a probability measure
$\mu_\psi : \qevents \rightarrow [0,1]$ on the space of events
defined as follows:
\[
\mu_\psi(A) = \ip{\psi}{A\psi}
\]
Similarlly to the classical case, this probability measure satisfies:
\begin{itemize}
\item $\mu(\mathbb{1}) = 1$, and
\item for all $A \in \qevents$, we have $\mu(A^*A) \geq 0$.
\end{itemize}

@@

Several assumptions are woven in the definition of a quantum probability space:
\begin{itemize}
\item the Hilbert space $\Hilb$;
\item the real interval $[0,1]$; 
\item the fact that each state induces a probability measure, i.e., the
Born rule~\cite{Born1984,Mermin2007}; 
\item the fact that every probability measure is induced by a state, i.e.,
Gleason's theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}. 
\end{itemize}
In the remainder of the paper, we examine each of these assumptions
and consider variations motivated by computation in a world with limited
resources. In particular, we will consider a variant of the Hilbert
space over finite fields~$\mathbb{F}_{p^{2}}^{d}$~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.
Instead of $\left[0,1\right]$, we will consider set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983}, in particular $\left\{ 0\right\} $,
impossible, and $\left(0,\infty\right)$, possible. Surprisingly,
some combinations of space and probability will result in no probability
measure or an unique probability measure. In these cases, there is
no need to discuss whether there is a Born rule, because we do not
have enough probability to correspond to every state.

If there may be more than one probability measure, we will discuss
whether there is a Born rule to generate a probability measure from
a state. When the space is $\mathbb{C}^{d}$, we will try to induced
a Born rule from the conventional Born rule; when the space is $\mathbb{F}_{p^{2}}^{d}$,
there is no natural way to induce a probability measure from a state,
so we will set some conditions a Born~$\tilde{\pi}$ should have: 
\begin{itemize}
\item Given a pure state $\ket{\Psi}\in\ffzd{p^{2}}$, a Born-rule~$\tilde{\pi}$
should give a probability~$\tilde{\pi}_{\Psi}$; 
\item $\ip{\Psi}{\Phi}=0\Leftrightarrow\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{0}$,
where $\tilde{0}$ is $0$ for $\left[0,1\right]$ and $\tilde{0}$
is impossible for $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $. 
\item $\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{\pi}_{\mathbf{U}\ket{\Psi}}\left(\mathbf{U}\ket{\Phi}\right)\textrm{ ,}$where
$\ket{\Psi},\ket{\Phi}\in\ffzd{p^{2}}$ and $\mathbf{U}$ is any unitary
map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbbm{1}$. 
\end{itemize}
Notice that when the space is $\mathbb{C}^{d}$, every Born rule we
consider will satisfy these three conditions.

@@

The results can be summarized in the following table:\\
\begin{tabular}{>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}}
\hline 
State space $\Hilb$ & Probability values & How many probability measure? & Is there a nature Born rule? & How many possible Born rule if there is no nature one? & How many probability measure not come from any possible Born rules?\tabularnewline
\hline 
\hline 
$\C^{2*}$ & $\left[0,1\right]$  & $\ge2$ & Yes &  & $\ge1$\tabularnewline
\hline 
$\C^{d*}$ for $d\ge3$ & $\left[0,1\right]$  & $\ge2$ & Yes &  & $0$\tabularnewline
\hline 
$\C^{d*}$ & $\mathscr{L}_{2}$ & $\ge2$ & Yes &  & $\ge1$\tabularnewline
\hline 
$\ff{3^{2}}^{d*}$ for $3\ge d\ge2$ & $\left[0,1\right]$  & $\ge2$ & No & $1$ & $\ge1$\tabularnewline
\hline 
$\ff{p^{2}}^{2*}$ for $p\ge7$ & $\left[0,1\right]$  & $\ge2$ & No & $\infty$ & $\ge1$\tabularnewline
\hline 
$\ff{7^{2}}^{3*}$ & $\left[0,1\right]$  & $1$ & No & $0$ & \tabularnewline
\hline 
$\ffzd{p^{2}}$for $d\ge3$ except $d=p=3$ & $\left[0,1\right]$  & $\ge1$ & No & $0$ & \tabularnewline
\hline 
$\ffzd{p^{2}}$ & $\mathscr{L}_{2}$ & $\ge2$ & No & $1$ & $\ge1$\tabularnewline
\hline 
\end{tabular}

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbm}
\usepackage{bbold}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{array}
\usepackage{mathrsfs}
\usepackage[all]{xy}

@@

We now formalize a \emph{quantum probability space} as
follows~\cite{Maassen2010,Swart2013}. We first assume an ambient
Hilbert space~$\Hilb$ and define the set of events $\qevents$ as all
linear combinations of \emph{projections} on~$\Hilb$. Specifically,
let $P_1, P_2, \ldots, P_k$ be mutually orthogonal projections on
$\Hilb$ with sum $\mathbb{1}$ and define the event space $\qevents$ to
be the linear span of these operators:
\[
\qevents=\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} .
\]
Each quantum state $\ket{\psi}$ induces a probability measure
$\mu_\psi : \qevents \rightarrow [0,1]$ on the space of events
defined as follows:
\[
\mu_\psi(A) = \ip{\psi}{A\psi}
\]
Similarly to the classical case, this probability measure satisfies: 
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for all $A\in\qevents$, we have $\mu(A^{*}A)\geq0$. 
\end{itemize}
\yutsung{The following is the diagram for quantum probability objects
and their corresponding physical meaning (in parentheses):
\[
\xymatrix{\textrm{projections (\textrm{events})}\ar[d]^{\mu_{\psi}\textrm{(probability)}} & \subsetneq & \textrm{Hermitian operators (observables)}\ar[d]^{\mu_{\psi}\textrm{(expectation value)}} & \subsetneq & \textrm{all operators in a }*\textrm{-algebra(?)}\ar[d]^{\mu_{\psi}\textrm{(?)}}\\
\left[0,1\right] & \subsetneq & \mathbb{R} & \subsetneq & \mathbb{C}
}
\]
Although the above paragraph claimed $\qevents$ is a set of events,
``all linear combinations of \emph{projections}'' and $\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} $
gives all operators in a $*$-algebra. Therefore, I got confused you
want $\qevents$ be ``the set of all events'' or ``the set of all
operators in a $*$-algebra''...}

@@

Several assumptions are woven in the definition of a quantum probability space:
\begin{itemize}
\item the Hilbert space $\Hilb$;
\item the real interval $[0,1]$; 
\item the fact that each state induces a probability measure, i.e., the
Born rule~\cite{Born1984,Mermin2007}; 
\item the fact that every probability measure is induced by a state, i.e.,
Gleason's theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}. 
\end{itemize}
In the remainder of the paper, we examine each of these assumptions
and consider variations motivated by computation in a world with limited
resources. In particular, we will consider a variant of the Hilbert
space over finite fields~$\mathbb{F}_{p^{2}}^{d}$~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.

For the probability values, we will consider set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983}, in particular $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $,
where impossible is a singleton set,$\left\{ 0\right\} $, and possible
is an open interval, $\left(0,\infty\right)$. Surprisingly, some
combinations of space and probability will result in an unique probability
measure. In these cases, there is no need to discuss whether there
is a Born rule, because we do not have enough probability to correspond
to every state.

If there may be more than one probability measure, we will discuss
whether there is a Born rule to generate a probability measure from
a state. When the space is $\mathbb{C}^{d}$, we will try to induced
a Born rule from the conventional Born rule. When the space is $\mathbb{F}_{p^{2}}^{d}$,
there is no natural way to induce a probability measure from a state,
so we will set some conditions for a Born rule~$\tilde{\pi}$: 
\begin{itemize}
\item Given a pure state $\ket{\Psi}\in\ffzd{p^{2}}$, a Born-rule~$\tilde{\pi}$
should give a probability~$\tilde{\pi}_{\Psi}$; 
\item $\ip{\Psi}{\Phi}=0\Leftrightarrow\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{0}$,
where $\tilde{0}$ is $0$ for $\left[0,1\right]$ and $\tilde{0}$
is impossible for $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $. 
\item $\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{\pi}_{\mathbf{U}\ket{\Psi}}\left(\mathbf{U}\ket{\Phi}\right)\textrm{ ,}$where
$\ket{\Psi},\ket{\Phi}\in\ffzd{p^{2}}$ and $\mathbf{U}$ is any unitary
map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbbm{1}$. 
\end{itemize}
Notice that when the space is $\mathbb{C}^{d}$, every Born rule we
consider will satisfy these three conditions.

@@

The results can be summarized in the following table:\\
\begin{tabular}{>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}}
\hline 
State space $\Hilb$ & Probability values & How many probability measure? & Is there a nature Born rule? & How many possible Born rule if there is no nature one? & How many probability measure not come from any possible Born rules?\tabularnewline
\hline 
\hline 
$\C^{2}$ & $\left[0,1\right]$  & $\ge2$ & Yes &  & $\ge1$\tabularnewline
\hline 
$\C^{d}$ for $d\ge3$ & $\left[0,1\right]$  & $\ge2$ & Yes &  & $0$\tabularnewline
\hline 
$\C^{d}$ & $\mathscr{L}_{2}$ & $\ge2$ & Yes &  & $\ge1$\tabularnewline
\hline 
$\ff{3^{2}}^{d}$ for $3\ge d\ge2$ & $\left[0,1\right]$  & $\ge2$ & No & $1$ & $\ge1$\tabularnewline
\hline 
$\ff{p^{2}}^{2}$ for $p\ge7$ & $\left[0,1\right]$  & $\ge2$ & No & $\infty$ & $\ge1$\tabularnewline
\hline 
$\ff{7^{2}}^{3}$ & $\left[0,1\right]$  & $1$ & No & $0$ & \tabularnewline
\hline 
$\ffzd{p^{2}}$for $d\ge3$ except $d=p=3$ & $\left[0,1\right]$  & $\ge1$

(Whether $\ge2$ or $1$?) & No & $0$ & \tabularnewline
\hline 
$\ffzd{p^{2}}$ & $\mathscr{L}_{2}$ & $\ge2$ & No & $1$ & $\ge1$\tabularnewline
\hline 
\end{tabular}
